---
title: "Lab 1: Intro to R"
author: "Al Haque"
date: "Jan 31,2022"
output: openintro::lab_report
---

```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
```

### Exercise 1

```{r view-girls-counts}
arbuthnot$girls
```


### Exercise 2

We can see that the girls were steadily getting baptized over the years with a sharp increase just after the 1660s and then there was a major decrease
right after the 1700s, but it immediately shot right back up the same year. 

```{r trend-girls}
ggplot(data=arbuthnot, aes(x=year,y=girls)) +
         geom_line()
```
```{r}
## I redefined the columns as the R markdown couldnt knit it with the object i.e object not found without defining these 
arbuthnot <- arbuthnot %>%
  mutate(total = boys + girls)
```






```{r}
## I redefined the columns as the R markdown couldnt knit it with the object i.e object not found without defining these 
arbuthnot <- arbuthnot %>%
  mutate(boy_ratio = boys/total)
```


### Exercise 3

This is the plot for the proportion of the boys over time. I see a lot of fluctuations in the graph as there were sharp increases and decreases to boys getting baptized notably there were a sharp increase in the years leading up to 1660s. 

```{r plot-prop-boys-arbuthnot}
ggplot(data=arbuthnot,aes(x=year,y=boy_ratio)) +
  geom_line()
```


### Exercise 4

We can see that there are 63 objects and 3 variables in this data frame. The column names in the present data set are year,boys,and girls. The years spanning from 1940 to 2002. 

```{r dim-present}
# Insert code for Exercise 4 here

data('present', package='openintro')
glimpse(present)
```


### Exercise 5

I made a new column called total for the present data frame and I compared the total values between arbuthnot data and the present data we can see that the present data frame is magnitudes greater than the arbuthnot data like the data has counts of over a million compared to the thousands of arbuthnot. 

```{r count-compare}
# Insert code for Exercise 5 here
present <-present %>%
  mutate(total=boys+girls)

## Total for arbuthnot
arbuthnot$total
```

```{r}
## Total for present
present$total
```


### Exercise 6

Arbuthnot's observation doesn't seem to hold up in the United States as it seems there has been a gradual decrease in the amount of boys being baptized. Al tough the ratio for boys is just above 50 percent. 

```{r plot-prop-boys-present}
present <- present %>%
  mutate(boy_ratio=boys/total)
```
```{r}
## Insert plot
library(ggplot2)
ggplot(data = present, aes(x=year,y=boy_ratio)) + 
  geom_line()

```
. 

### Exercise 7

The year we saw the most births in the US is 1961 with the total being 4268326.

```{r find-max-total}
present %>%
  arrange(desc(total))
```

